---
title: "R Notebook"
output:
  html_document:
    keep_md: yes
    self_contained: no
---
This is an exploration of the John Hopkins data.

```{r}
library(readr)
library(data.table)
library(ggplot2)
library(plotly)
library(directlabels)
library(dplyr)

```



```{r}

confirmed_cases<-data.table(read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))
country_populations<-data.table(read.csv("population-figures-by-country-csv_csv.csv"))

ma_lag_length<-3

```

Let's get cases by country, to make this a bit easier.

```{r}
confirmed_cases<-confirmed_cases[,1:(ncol(confirmed_cases)-1)]
transform_cols<-colnames(confirmed_cases)[5:ncol(confirmed_cases)]
id_vars <- c("Country/Region","Province/State","Lat","Long")


confirmed_cases_long <- melt(confirmed_cases,
                             id.vars=id_vars,
                             measure.vars = transform_cols,
                             variable.name = "case_date",
                             value.name = "Confirmed Cases"
                             )

confirmed_cases_long_bycountry <- confirmed_cases_long[,.(`Confirmed Cases`=sum(`Confirmed Cases`)),by=.(`Country/Region`,case_date)]

pops_clean<-country_populations[,.(Country,Population2016=Year_2016)]
pops_clean<-rbind(pops_clean,data.table("Country"="Taiwan*","Population2016"=23800000))
population_renaming<-list(list("Brunei","Brunei Darussalam"),
                          list("Czechia","Czech Republic"),
                          list("Egypt","Egypt, Arab Rep."),
                          list("French Guiana","Guyana"),
                          list("Iran","Iran, Islamic Rep."),
                          list("Korea, South","Korea, Rep."),
                          list("North Macedonia","Macedonia, FYR"),
                          list("Russia","Russian Federation"),
                          list("Slovakia","Slovak Republic"),
                          list("US","United States"))
for (poptuple in population_renaming){
  pops_clean[Country==poptuple[[2]],Country:=poptuple[[1]]]
}


confirmed_cases_long_bycountry<-
  merge(confirmed_cases_long_bycountry,
        pops_clean,
        by.x="Country/Region",
        by.y="Country",
        all.x=TRUE
        )
#View(confirmed_cases_long_bycountry[,.(Pop=mean(Population2016)),by=`Country/Region`] %>% .[is.na(Pop)])
```





```{r}

confirmed_cases_long_bycountry$case_date<-as.Date(confirmed_cases_long_bycountry$case_date,format="%m/%d/%y")

threshold<-max(confirmed_cases_long_bycountry[`Country/Region`=="New Zealand",`Confirmed Cases`])
#mark which days have reached the threshold
confirmed_cases_long_bycountry[,IsAtOrPastThreshold:=(`Confirmed Cases`>=threshold)]

#identify the first day for each country that has reached the threshold
confirmed_cases_long_bycountry[,
                               DatePastThreshold:=min(.SD[IsAtOrPastThreshold==TRUE,case_date]),
                               by=`Country/Region`]

#then calculate the case_date relative to the threshold
confirmed_cases_long_bycountry[,DaysPastThreshold:=as.integer(case_date-DatePastThreshold)]


confirmed_cases_long_bycountry$case_date<-as.Date(confirmed_cases_long_bycountry$case_date,format="%m/%d/%y")

     #make it easy to see the US & UK
     confirmed_cases_long_bycountry[`Country/Region`=="US",
                                    `Country/Region`:="United States",
                                    ]
     confirmed_cases_long_bycountry[`Country/Region`=="UK",
                                    `Country/Region`:="United Kingdom",
                                    ]
```


```{r}



#mark which days have reached the threshold
confirmed_cases_long_bycountry[,IsAtOrPastThreshold:=(`Confirmed Cases`>=threshold)]

#identify the first day for each country that has reached the threshold
confirmed_cases_long_bycountry[,
                              DatePastThreshold:=min(.SD[IsAtOrPastThreshold==TRUE,case_date]),
                              by=`Country/Region`]

#then calculate the date relative to the threshold
confirmed_cases_long_bycountry[,DaysPastThreshold:=as.integer(case_date-DatePastThreshold)]

confirmed_graph<-confirmed_cases_long_bycountry
#Remove China if it was above the threshold on the first day. we don't do this to any other countries
#but China is further along than all others.
if(threshold<min(confirmed_graph[`Country/Region`=="China",`Confirmed Cases`])){
 confirmed_graph <- 
   confirmed_graph[`Country/Region`!="China"]
}

#now - set a 'target country'
target_country <- "New Zealand"
#we look to see which is closet to that country at that threshold and in the days prior
target_c_data<-confirmed_graph[
  `Country/Region`=="New Zealand",
  .(TargetConfirmedCases=`Confirmed Cases`,DaysPastThreshold)]


target_c_data$DaysPastThreshold

confirmed_graph<-merge(
  confirmed_graph,
  target_c_data,all.x=TRUE)

#now for each country
#which has the smallest sum of squares difference from NZ's progression?
countries_diff<-confirmed_graph[DaysPastThreshold>-7 & DaysPastThreshold<=0
                  ,.(SumSquaresDiff=sum((TargetConfirmedCases-`Confirmed Cases`)^2))
                  ,by=`Country/Region`]

comparison_countries<-as.character(
  countries_diff[order(SumSquaresDiff)] %>% 
    .[2:6,`Country/Region`]
  )


```


```{r}
case_limit<-20
days_limit <- 21





gg_confirmed_basic<-function(confirmed_graph,comparison_countries){
  #define colour groups
  confirmed_graph$SelectorStatus<-"DUnselected"
  confirmed_graph[`Country/Region`==target_country, SelectorStatus:="CTarget"]
  confirmed_graph[`Country/Region` %in%comparison_countries , SelectorStatus:="BComparators"]
  other_countries<-setdiff(unique(confirmed_graph$`Country/Region`),c(comparison_countries,target_country))
  confirmed_graph$`Country/Region`<-factor(
    confirmed_graph$`Country/Region`,
    levels=c(other_countries,comparison_countries,target_country),
    order=TRUE
  )
  # confirmed_graph$SelectorStatus<-factor(
  #   confirmed_graph$SelectorStatus,
  #   levels=c("CTarget","BComparators","DUnselected"),
  #   order=TRUE
  #                                           
  #                                           )
    #now we want a list of the lines
  #with an x axis point either at the end of the graph,
  #or at the end of the line, whichever is first,     
  country_labels<-confirmed_graph[#`Country/Region` %in% c(comparison_countries,target_country)
    ,
                                .(intercept_day=min(days_limit,max(DaysPastThreshold))),
                                by=`Country/Region`] %>% .[!is.na(intercept_day)] %>%
   #and then at that x-axis point, we want the y-axis point.  
   merge(confirmed_graph[,.(`Country/Region`,DaysPastThreshold,`Confirmed Cases`,
                            DatePastThreshold,SelectorStatus)],
                       by.x=c("Country/Region","intercept_day"),
                       by.y=c("Country/Region","DaysPastThreshold"),
         all.x=TRUE,all.y=FALSE)
  
  x_breaks  <- seq(-7,days_limit,7)
  x_labels <- c(paste(abs(x_breaks[1]),"Days Before"),
               x_breaks[2:(length(x_breaks)-1)],
               paste(x_breaks[length(x_breaks)],"Days After"))

  ggplot(
      confirmed_graph[order(SelectorStatus),],
      aes(x=DaysPastThreshold,
          y=`Confirmed Cases`,
          group=`Country/Region`,
          color=SelectorStatus,#DatePastThreshold
          alpha=SelectorStatus
            ))+
 geom_line(size=1)+
 #geom_point(alpha=0.5)+
 scale_x_continuous(limits=c(-7,days_limit),
                    breaks=x_breaks,
                    labels = x_labels)+
 coord_cartesian(xlim=c(-7,days_limit+3),
                 ylim=c(20,max(confirmed_graph$`Confirmed Cases`)))+
 scale_y_continuous(trans = "log10",
                    breaks = c(1,2,5,10,
                               20,50,100,200,500,1000,2000,5000,10000,20000,50000,100000),
                    minor_breaks=NULL,
                    labels = scales::comma,
                    sec.axis = dup_axis())+
 geom_vline(xintercept = 0,linetype="dashed")+
  scale_color_manual(values=c("#3333ff","#aa0000","#222222"))+
  scale_alpha_manual(values=c(1,1,0.2))+#c(0.2,1,1))+
    theme(legend.position = "none")+
 geom_text(data=country_labels,
           aes(x=intercept_day,y=`Confirmed Cases`,label=`Country/Region`),
           hjust=0,size=3,fontface="bold")
  }
  
myplot<-gg_confirmed_basic(confirmed_graph,comparison_countries)+
  labs(title=paste0("Growth in cases by day from the ",threshold," case mark, by country"),
      x=paste0("Days past first day with ",threshold," cases"),
      y="Total confirmed cases",
      color="Date Country Reached Threshold")
 


myplot
```



```{r}
case_limit<-20
days_limit <- 21


myplot<-gg_confirmed_basic(confirmed_graph,
                           c()
                           )+labs(
  title="We can compare New Zealand's progression with locations around the world from the same time",
      x=paste0("Days past first day with ",threshold," cases"),
      y="Total confirmed cases",
      color="Date Country Reached Threshold")
 


myplot

```


```{r}
case_limit<-20
days_limit <- 21


myplot<-gg_confirmed_basic(confirmed_graph,
                           c("Singapore","Spain","Italy")
                           )+labs(
  title="There's a world of difference between the best cases, Singapore,\nand the worst cases, Italy and Spain",
      x=paste0("Days past first day with ",threshold," cases"),
      y="Total confirmed cases",
      color="Date Country Reached Threshold")
 


myplot

```




```{r}
case_limit<-20
days_limit <- 21


myplot<-gg_confirmed_basic(confirmed_graph,
                          comparison_countries
                           )+labs(
  title="New Zealand following along a path\npreviously led by the Panama, Luxembourg, Ireland, Austria, and France",
      x=paste0("Days past first day with ",threshold," cases"),
      y="Total confirmed cases",
      color="Date Country Reached Threshold",
  caption=" Source: John Hopkins Dataset.\n http://github.com/bjsmith @curiouskiwicat benjsmith.shinyapps.io/john_hopkins/")
 


myplot

```

Which are most similar to us?

Finland is now up to 450 cases with community transmission. 7 days onward (16 March), they implemented a raft of severe measures, including banning gatherings of more than 10 people and closing down schools from primary level up.

https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_Finland

Chile reached 43 cases on 13 March. On the same day, they banned events of more than 500 people. The borders were closed three days later, and soon after, a "State of Catastrophe" was declared. Schools were shut down, and public gatherings were limited.

Austria reached 41 cases on 5 March. Universities, schools, and restaurants were closed by 17 March. Currently, there are over 2600 cases and 6 fatalities.
